\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-03-06 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
section 1

\medskip
In large scale field operations, fingering is a relatively common condition in reservoirs with water-injection wells. As water and oil interacts, the interface between those two fluids moves along the geological formation, creating an uneven, or fingered, profile. The result of fingering is an inefficient sweeping action that can bypass significant volumes of recoverable oil and, in severe cases, an early breakthrough of water into adjacent production wellbores. This instability is mainly governed by the ratio of mobility of the displacing fluid over that of the displaced fluid (Mobility-ratio, MR). %In a water-oil system the MR is given by eq.\ref{eq:MR}.
However, other important factors also influence the severity of viscous fingering: heterogeneous permeability fields, gravity, anisotropic dispersion, velocity dependence of dispersion, nonmonotonic viscosity profiles, etc. High mobility-ratio displacement processes introduces very complex fingering patterns.

@@

\medskip 
The conservation of mass along with the Darcy's law are the governing equations that describe the flow dynamics in porous media. Several computational methods have been developed to accurately represent force balanceand mass conservation in darcean flows.
Finite difference methods (FDM) have been extensively used in most industry-standard flow simulators~\citep{aziz_1986, chen_2005, chang_1990} with relative success in multi-physics single phase applications (e.g., black oil model, passive tracer advection etc). A finite difference method (FDM) discretization is based upon the differential form of the PDE to be solved. 
%Analytic models can be used to analyze reservoir performance as long as the reservoir is not highly heterogeneous. However, FDM are often limited to relatively simple geometries (with low grid quality and orientation) representing idealised geological formations~\citep{mlacnik_2004, king_1999}. Additionally, FDM produce excessive numerical dispersion when heterogeneity (represented by permeability and porosity fields) is present~\citep{chavent_1986}.
The geometrical flexibility associated with high-order numerical accuracy of finite element methods (FEM) has proven to overcome these deficiencies. A finite element method (FEM) discretization is based upon a piecewise representation of the solution in terms of specified basis functions. The computational domain is divided up into smaller domains (finite elements) and the solution in each element is constructed from the basis functions. Advantages of FEM are the better description of the geometry and complex shapes, better treatment of fluid flow and saccurate more accurate stress calculations using a smaller number of nodes, less memory space and shorter run times. At this point let us remind that a comparison of computaional methods is difficult and depends always on the test-case.
%Even though the most popular ones for the fluid flow equations are the Finite Difference Methods (FDM) there are also the Finite Volume Methods (FVM) and the Finite Element Methods. The latter is the one that will be used in this paper and more precisely a member of the FEM family, the Control-Volume FEM method (CVFEM).

@@

%~\citep{forsyth_1990, geiger_2004, hurtado_2007}. 
%In this formulation, standard Galerkin-FEM method is used to discretise the pressure equation (FEM representation of pressure and velocity fields) whereas a control volume-based formulation is used to discretise all remaining scalar fields~\citep[e.g., saturation, density, species concentration etc, see][]{gomes_book_2012}.

@@

\medskip
In the CVFE method pressure is interpollated using piecewise linear finite element basis functions, also known as shape functions, %$\underline{\bf v}$ and $\underline{\bf p}$) 
while saturation is represented in node centered CV basis functions respectevely, and the saturation equation is solved explicitly after solving for pressure at each time.

>>>>

\medskip
In large scale field operations, fingering is a relatively common condition in reservoirs with water-injection wells. As water and oil interacts, the interface between those two fluids moves along the geological formation, creating an uneven, or fingered, profile. The result of fingering is an inefficient sweeping action that can bypass significant volumes of recoverable oil and, in severe cases, an early breakthrough of water into adjacent production wellbores. This instability is mainly governed by the ratio of mobility of the displacing fluid and the of the displaced fluid (mobility-ratio, MR). %In a water-oil system the MR is given by eq.\ref{eq:MR}.
However, other important factors also influence the severity of viscous fingering: heterogeneous permeability fields, gravitational forces, anisotropic dispersion, velocity dependence of dispersion, non-monotonic viscosity profiles, etc.

@@

\medskip 
The conservation of mass along with the Darcy's law are the governing equations that describe the flow dynamics in porous media. Several computational methods have been developed to accurately represent force balanceand mass conservation in Darcean flows.
Finite difference methods (FDM) have been extensively used in most industry-standard flow simulators~\citep{aziz_1986, chen_2005, chang_1990} with relative success in multi-physics single phase applications (e.g., black oil model, passive tracer advection etc). 
%The FDM discretization is based upon the differential form of the PDE to be solved. 
%Analytic models can be used to analyze reservoir performance as long as the reservoir is not highly heterogeneous. However, FDM are often limited to relatively simple geometries (with low grid quality and orientation) representing idealised geological formations~\citep{mlacnik_2004, king_1999}. Additionally, FDM produce excessive numerical dispersion when heterogeneity (represented by permeability and porosity fields) is present~\citep{chavent_1986}.
The geometrical flexibility associated with high-order numerical accuracy of finite element methods (FEM) has proven to be more efficient compared with FDM schemes when they are used to solve complex geometries. A finite element method (FEM) discretization is based upon a piecewise representation of the solution in terms of specified basis functions. The computational domain is divided up into smaller domains (finite elements) and the solution in each element is constructed from the basis functions. Advantages of FEM are the better description of the geometry and complex shapes, better treatment of fluid flow and saccurate more accurate stress calculations using a smaller number of nodes, less memory space and shorter run times. %At this point let us remind that a comparison of computaional methods is difficult and depends always on the test-case.
%Even though the most popular ones for the fluid flow equations are the Finite Difference Methods (FDM) there are also the Finite Volume Methods (FVM) and the Finite Element Methods. The latter is the one that will be used in this paper and more precisely a member of the FEM family, the Control-Volume FEM method (CVFEM).

@@

@@

In the CVFEM pressure is interpollated using piecewise linear finite element basis functions, %$\underline{\bf v}$ and $\underline{\bf p}$) 
while saturation is represented in node centered CV basis functions respectevely, and the saturation equation is solved explicitly after solving for pressure at each time.
